Adopting agile in the large: motivations, success factors and challenges

### 2014-05-03 journal/article.tex
Restructuring: Change resistance, other organizational problems

\subsubsection{Tensions between management and development}

% In the beginning of the transformation the organization has realized the need
% to change and is going towards the initial push.
In many cases the initiative to change came from management, but when presented
wrong the initiative becomes a mandate. Spayd summarizes this as:
``Organizations do not change merely because the boss says so, at least not in
the way that is intended'' [P49]. In that organization the introduction of CMM
was carried out consistently by a mandate, but the collaborative nature of agile
did not comply with such a mindset. A top down mandate may also dilute the
understanding of the reasons for starting the transformation, and the
understanding of agile development overall [P2]. O'Connor describes how team
members became sceptical towards agile when implementing mandaetd changes did
not bring any visible benefits [P37].

@@

The problem of tension between management and development may also be opposite.
In cases where the change was emerging bottom up the reluctancy to change was
not in the team members but management.
This caused that significant organizational change could not be raised above the
team level. For instance, Atlas describes how it would have required executive
support to integrate the product management office and separate quality
assurance groups in the agile process [P3]. In this case, when managers owning
projects with many teams were not involved it retrained agile development to
spread beyond team level. Further, if management does not speak out a goal
towards using agile, developers feel that the agile methods may be replaced by
something else at any time [P12].
Lack middle management support was seen as the most serious shortcoming by Spayd
[P49]. Middle management is in a position to undermine the entire
transformation, and may do so if they do not participate in, and thus
understand, the agile method. Agile brings changes to some management roles
[P2], and lack of understanding of agile development will leave managers feeling
left out [P6].

@@

A condition hindering the transformation was that old commitments still needed
to be tended to even though the organization was being restructured.
Some organizations started their agile journey in a state where people were
overcommitted, but it was later realized that overloaded people will not be able
to learn new ways of working and changing their behavior [P37, P42]. Cowan
describes how time was allocated for tending to old commitments [P11]. However,
even when extra work was prepared for people with specific expertise
became overloaded. After time it was realized that overcommitment must be
reduced by allowing to shift excess work for later time [P11]. In one case
people were forced to remain committed to firm deadlines which resulted in
ignoring new practices introduced by agile, and eventually in the agile method
breaking down.

% ADD THIS QUOTATION:
% [P21]
% The development projects require continues maintenance and support. In addition,
% the team was working on a new project with firm deadlines. The work environment
% was very demanding and team worked under pressure to produce products according
% to the planned schedule.
% Scrum adoption process started while every member of the team was engaged in
% his/her everyday tasks. With such work pressure the daily Scrum meetings added
% extra pressure to the employees. They used to meet weekly and later twice a
% month and then only when required and usually after working hours. As teams
% started to skip daily meetings it also affected the learning process of scrum
% between the team members. This eventually led to the failure of learn and
% implement concept of the agile method.

@@

Change resistance is a central challenge in any organizational transformation.
People will not be willing to change unless there are good reasons, and the
change is preceived easy enough. It is difficult to attain a buy-in for a
change, and even organizations that have a flexible culture will face resistance
[P10]. It should be expected that not everyone will be willing to change, and
some employees will never adapt to the change [P6]. In one case it was reported
that objections from team members was a major reason for loss in time and
productivity [P29].

\subsubsection{Lack of resources}

Lack of resources allocated for supporting the transformation was an evident
problem in many cases. Resources were most often short for training and
coaching, but underbudgeting these activities may be a serious mistake.
Hajjdiab [P20] describes how reluctance of management to invest in training
caused teams to be ill prepared for the agile transformation effort, and
eventually resulted in ending the use of agile methods. As a less dramatic
outcome of shortened training was lower motivation [P45].

>>>>

To start a transformation the organization needs an initial push. However,
people are usually disincluned to change, and it is likely that resistance
will emerge in any change. 
% Change resistance is a central challenge in any organizational transformation.
People will not be willing to change unless there are good reasons, and the
change is preceived easy enough. It is difficult to attain a buy-in for a
change, and even organizations that have a flexible culture will face resistance
[P10]. It should be expected that not everyone will be willing to change, and
some employees will never adapt to the change [P6]. In one case it was reported
that objections from team members was a major reason for loss in time and
productivity [P29].

The way the transformation is initiated will also affect how change resistance
will show.
In many cases the initiative to change came from management, but when presented
wrong the initiative becomes a mandate that pepole are not receptive for. Spayd
[P49] summarizes this as: ``Organizations do not change merely because the boss
says so, at least not in the way that is intended''. In that organization the
introduction of CMM had been carried out consistently by a mandate, but the
collaborative nature of agile did not mix well with such a mindset. A top down
mandate may also dilute the understanding of the reasons for starting the
transformation, and the understanding of agile development overall [P2].
O'Connor describes how team members became sceptical towards agile when
implementing mandaetd changes did not bring any visible benefits [P37].

@@

The problem of tension between management and development may also be opposite.
In cases where the change was emerging bottom up the reluctancy to change was
not in the team members but management. This caused that significant
organizational change could not be raised above the team level.
For instance, Atlas [P3] describes how it would have required executive support
to integrate the product management office and separate quality assurance groups
in the agile process. In this case, when managers owning projects with many
teams were not involved in the transformation, the agile model was restrained
from spreading beyond team level. Lack of middle management support was also
seen as the most serious shortcoming by Spayd [P49]. Middle management is in a
position to undermine the entire transformation, and may do so if they do not
participate in, and thus understand, the agile method. Agile brings changes to
some management roles [P2], and lack of understanding of agile development will
leave managers feeling left out [P6]. Still one problem is that if management
does not speak out a goal towards using agile, developers feel that the agile
methods may be replaced by something else at any time [P12].

@@

A condition hindering the transformation was that old commitments still needed
to be tended to even though the organization was being restructured.
Some organizations started their agile journey in a state where people were
overcommitted, but it was later realized that overloaded people will not be able
to learn new ways of working and changing their behavior [P37, P42]. Cowan [P11]
describes how time was allocated for tending to old commitments. However, even
when extra work was prepared for people with specific expertise became
overloaded. After time it was realized that overcommitment must be reduced by
allowing to shift excess work for later time [P11]. In another case management
forced people to remain committed to firm deadlines, even midst in the
organizational transformation. The engagement in old commitments and tasks
resulted in ignoring new practices introduced by agile, and eventually the agile
method was breaking down [P21].

@@

% Lack of resources

Lack of resources allocated for supporting a transformation was an evident
problem. Resources were most often short for training and coaching, but
underbudgeting these activities may be a serious mistake.
Hajjdiab [P20] describes how reluctance of management to invest in training
caused teams to be ill prepared for the agile transformation effort, and
eventually resulted in ending the use of agile methods. As a less dramatic
outcome of shortened training was lower motivation [P45].
